
---
title: "Decoding Intracellular Pathogen of H3N2 at the Single-Cell level 
  using Yeskit"
author: "Wei Zhang"
date: "2021/09/10"
output:
  html_document:
    toc: yes
    toc_depth: 4
    toc_float: no
    theme: cerulean
    highlight: tango
    number_sections: yes
    df_print: tibble
  pdf_document:
    toc: yes
    toc_depth: '4'
    fig_caption: yes
vignette: >
  %\VignetteIndexEntry{Tutorial}
  %\usepackage[UTF-8]{inputenc}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
---

```{r, echo=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = FALSE, eval = TRUE,
                      warnings = FALSE, message = FALSE,
                      fig.width = 6, fig.height = 5)
```

Taking the in-vitro experiment of H3N2 infection data (SRA Accession number: 
  [SRP239555](https://www.ncbi.nlm.nih.gov/sra/?term=SRP239555)) as an 
  example, we used [PathogenTrack](https://github.com/ncrna/PathogenTrack) to 
  identify H3N2 infected cells at the single-cell level and used 
  [Yeskit](https://github.com/ncrna/Yeskit) to analyze and explore the 
  biological functions that may be related to H3N2 infection.

# Install Yeskit from GitHub
```{r installation, eval=FALSE}
if (!requireNamespace("BiocManager", quietly = TRUE))
    install.packages("BiocManager")
if (!requireNamespace("devtools", quietly = TRUE))
    BiocManager::install("devtools")
if (requireNamespace("Yeskit", quietly = TRUE))
    devtools::install_github("ncrna/Yeskit")
```

# Import Yeskit
First, we load the package:
```{r setup, warning=FALSE, message=FALSE}
library(Yeskit)
library(topGO)
```

# Importation
Now, let's load the single-cell count matrix:
```{r import-bystander-data, fig.width=5, fig.height=5, fig.align="center"}
#| fig.cap="Quality control for bystander"
Bystander <- scRead(sample_name = "Bystander", 
                    data_dir = system.file(
                      "extdata/H3N2_10X_matrix/Bystander/", 
                      package="Yeskit"), 
                    gene_column = 2, project_name = "H3N2", 
                    group_name = "Bystander", 
                    meta_file = system.file(
                      "extdata/H3N2_10X_matrix/Bystander/microbes.tsv", 
                      package="Yeskit")
                    )
```

```{r import-infected-data, fig.width=5, fig.height=5, fig.align="center"}
#| fig.cap="Quality control for bystander"
Infected <- scRead(sample_name = "Infected", 
                   data_dir = system.file(
                     "extdata/H3N2_10X_matrix/Infected/", package="Yeskit"), 
                   gene_column = 2, project_name = "H3N2", 
                   group_name = "Infected", 
                   meta_file = system.file(
                     "extdata/H3N2_10X_matrix/Infected/microbes.tsv", 
                     package="Yeskit"))
```

# Integration
Then, we integrate these two `Seurat` object
```{r integration, message=FALSE, warning=FALSE, fig.align="center"}
#| fig.cap="Harmony Interation"
Integrated <- scIntegrate(object.list=list(Bystander, Infected), 
                          object.names = c("Bystander", "Infected"), 
                          batch.rm = "harmony", resolution = 0.7)
```

# Differential analysis
## Differential analysis between Infected and Bystander
```{r DGE-of-infected-vs-bystander, message=FALSE, warning=FALSE}
slot(Integrated, "misc")$Infected_vs_Bystander <- scDGE(object = Integrated, 
                         comparison = c("Infected", "Bystander"), 
                         group.by = "group", min.cells = 10, 
                         logFC = 0.25, clusters = NULL)
```

## Differential analysis between H3N2_positive and H3N2_negative 
```{r DGE-of-H3N2pos-vs-H3N2neg, message=FALSE, warning=FALSE}
slot(Integrated, "misc")$H3N2 <- scPathogenDGE(object = Integrated, 
                         species.by = "H3N2", min.cells = 5)
```

# GO annotation
## GO annotation of DGEs between Infected and Bystander
```{r infected-vs-bystander-GO-Enrichment, message=FALSE, warning=FALSE}
slot(Integrated, "misc")$Infected_vs_Bystander.GO <- scGO(object = Integrated, 
                         key = "Infected_vs_Bystander", 
                         logFC = 0.25, only.pos = FALSE,
                         reference = "human")
```

## GO annotation of DGEs between H3N2_positive and H3N2_negative 
```{r H3N2pos-vs-H3N2neg-GO-Enrichment, message=FALSE, warning=FALSE}
slot(Integrated, "misc")$H3N2.GO <- scPathogenGO(object = Integrated, 
                         key = "H3N2", clusters = NULL, species = "H3N2", 
                         logFC = 0.25)
```

# MSigDB scoring
```{r MSigDB-score, message=FALSE, warning=FALSE}
Integrated <- scMsigdbScoring(object = Integrated, 
                              category = "H", 
                              genesets = "HALLMARK_INFLAMMATORY_RESPONSE"
                              )
```

# Visualization
## Visualization of cell clusters by scDimPlot
```{r DimPlot, fig.height=3, fig.width=8, fig.align="center"}
#| fig.cap="Cell DimPlot"
scDimPlot(object = Integrated, 
          reduction = "umap", 
          cols = NULL, 
          split.by = "sample", 
          ncol = 2, 
          pt.size = 2)
```

## Visualization of cell densities by scDensityPlot
```{r DensityPlot, fig.height=3, fig.width=8, fig.align="center"}
#| fig.cap="Cell Density Plot"
scDensityPlot(object = Integrated, 
              reduction = "umap", 
              split.by = "sample", 
              ncol = 2)
```

## Visualization of cell population fractions by scPopulationPlot, 
  the x axis stands for clusters
```{r PopulationPlot, fig.height=3, fig.width=3, fig.align="center"}
#| fig.cap="Cell Population Plot by cluster"
scPopulationPlot(object = Integrated, 
                 by = "cluster", 
                 cols = "sc", 
                 order = c("Bystander", "Infected"))
```

## Visualization of meta data by scVizMeta
```{r VizMetaPlot, fig.height=3.3, fig.width=10, fig.align="center"} 
#| fig.cap="H3N2 DimPlot"
scVizMeta(object = Integrated, 
          reduction = "umap", 
          signature="H3N2", 
          title = "H3N2", 
          raster = TRUE, 
          split.by = "sample", 
          pt.size = 2, 
          interval = c(
            Abundant = 1000, 
            Large = 500, 
            Medium = 100, 
            Small = 10, 
            Single = 1, 
            None = 0)
          )
```

## Visualization of H3N2-infected cell fractions by scPathogenRatioPlot
```{r PathogenRatioPlot, fig.height=3, fig.width=4, fig.align="center"}
#| fig.cap="H3N2 Ratio Plot"
scPathogenRatioPlot(object = Integrated, 
                    species = "H3N2", 
                    split.by = "sample", 
                    ncol = 2)
```

## Visualization of DGEs by scVolcanoPlot
```{r VolcanoPlot, fig.height=4, fig.width=4, fig.align="center"}
#| fig.cap="Volcano Plot of Infected_vs_Bystander"
scVolcanoPlot(object = Integrated, 
              key = "Infected_vs_Bystander", 
              cluster = "0", 
              top_n = 10)
```

## Visualization of enriched GO terms for up-regulated genes by scGOBarPlot
```{r GOBarPlot-up, fig.height=3, fig.width=6, fig.align="center"}
#| fig.cap="Infected_vs_Bystander.GO_up"
scGOBarPlot(object = Integrated, 
            key = "Infected_vs_Bystander.GO", 
            ont = "BP", 
            top_n = 6, 
            direction = "up", 
            cluster = "0")
```

## Visualization of enriched GO terms for down-regulated genes by scGOBarPlot
```{r GOBarPlot-down, fig.height=3, fig.width=6, fig.align="center"}
#| fig.cap="Infected_vs_Bystander.GO_down"
scGOBarPlot(object = Integrated, 
            key = "Infected_vs_Bystander.GO", 
            ont = "BP", 
            top_n = 6, 
            direction = "down", 
            cluster = "0")
```

## Visualization of enriched GO terms for up-regulated genes by scGODotPlot
```{r GODotPlot-up, fig.height=4, fig.width=7, fig.align="center"}
#| fig.cap="Infected_vs_Bystander.GO_up"
scGODotPlot(object = Integrated, 
            key = "Infected_vs_Bystander.GO", 
            ont = "BP", 
            direction = "up", 
            top_n = 10, 
            font.size = 8)
```

## Visualization of enriched GO terms for down-regulated genes by scGODotPlot
```{r GODotPlot-down, fig.height=4, fig.width=7, fig.align="center"}
#| fig.cap="Infected_vs_Bystander.GO_down"
scGODotPlot(object = Integrated, 
            key = "Infected_vs_Bystander.GO", 
            ont = "BP", 
            direction = "down", 
            top_n = 10,
            clusters = c("0"),
            font.size = 8)
```


## BarPlot shows enriched GO terms for up-regulated genes of H3N2pos_vs_H3N2neg
```{r GOBarPlot-H3N2-up, fig.height=4, fig.width=7, fig.align="center"}
#| fig.cap="H3N2pos_vs_H3N2neg.GO_up"
scGOBarPlot(object = Integrated, 
            key = "H3N2.GO", 
            cluster = "0", 
            extra = "H3N2", 
            top_n = 6)
```

## DotPlot shows enriched GO terms for up-regulated genes of H3N2pos_vs_H3N2neg
```{r GODotPlot-H3N2-up, fig.height=4, fig.width=7, fig.align="center"}
#| fig.cap="H3N2pos_vs_H3N2neg.GO_up"
scGODotPlot(object = Integrated, 
            key = "H3N2.GO", 
            clusters = "0", 
            extra = "H3N2", 
            top_n = 6)
```

## Visualization of HALLMARK_INFLAMMATORY_RESPONSE pathway
```{r INFLAMMATORY-RESPONSE, fig.height=3, fig.width=9, fig.align="center"}
#| fig.cap = "HALLMARK_INFLAMMATORY_RESPONSE"
scScoreDimPlot(object = Integrated, 
               signature = "HALLMARK_INFLAMMATORY_RESPONSE", 
               split.by="sample", 
               ncol = 2, 
               pt.size = 2)
```

# Session Information
```{r, echo=FALSE}
sessionInfo()
```

# References

1. Zheng GXY, Terry JM, Belgrader P et al. Massively parallel digital 
  transcriptional profiling of single cells, Nature Communications 2017;8:14049
2. Stuart T, Butler A, Hoffman P et al. Comprehensive Integration of 
  Single-Cell Data, Cell 2019;177:1888-1902.e1821
3. Korsunsky I, Millard N, Fan J et al. Fast, sensitive and accurate int-
  egration of single-cell data with Harmony, Nature Methods 2019;16:1289-1296
4. Alexa A, Rahnenführer J. Gene set enrichment analysis with topGO, 
  Bioconductor Improv 2009;27:1-26
5. Liberzon A, Subramanian A, Pinchback R et al. Molecular signatures database 
  (MSigDB) 3.0, Bioinformatics 2011;27:1739-1740
6. Yuan et al. Computational Methods for Single-Cell Data Analysis. Methods in 
  Molecular Biology. 2019;doi:10.1007/978-1-4939-9057-3
7. Borcherding N, Bormann N L, Kraus G. scRepertoire: An R-based toolkit for 
  single-cell immune receptor analysis. F1000Research, 2020, 9
8. Petukhov V. ggrastr: Raster layers for ggplot2. 2018
9. Wickham H. ggplot2. Wiley Interdisciplinary Reviews: Computational 
  Statistics, 2011, 3(2): 180-185
10. Finak G, McDavid A, Yajima M, et al. MAST: a flexible statistical 
  framework for assessing transcriptional changes and characterizing 
  heterogeneity in single-cell RNA sequencing data. 
  Genome biology, 2015, 16(1): 1-13.
